RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-31 casestudies.tex
yes

We present an example Stock Price Analysis application
in RCloud.\footnote{Unfortunately,
we could not include production notebooks containing proprietary or personal
information.} The example, although simplified, shows key steps in the
development and deployment process we aimed to support.

@@

The example shows a sequence of visualizations of the performance
of financial stocks over a multi-year period. The first visualization
in this example uses ggplot2 and was written by Hadley Wickham.
It reads data provided by Yahoo! Finance as a web service, and shows
the price for a single trading symbol.

>>>>

To illustrate some aspects of RCloud, we present an example application
for stock price analysis.\footnote{Unfortunately we could not include
production notebooks containing proprietary or personal information.}
The example, although simplified, shows key steps in the development
and deployment process we intend to support.

@@

It shows a sequence of visualizations of the performance
of financial stocks over a multi-year period. The first visualization
in this example uses ggplot2 and was written by Hadley Wickham.
It reads data provided by Yahoo! Finance as a web service, and shows
the price for a single trading symbol.

### 2014-03-31 system.tex
yes

As a result, HTTP is the lingua franca of interprocess communication
(IPC). One of the design goals for RCloud was to provide an attractive
environment for creators of data-analysis scripts, that also behaves
as a first-class citizen in the pre-existing ecosystem of computer
services and networks with an organization.

@@

One of the main differences between RCloud and other systems is the
notion that notebooks are ``always deployed''. This has a variety of
advantages which we discuss \todo{where?}. On the other hand, there
are a variety of situations that might ask for a specific version of a
notebook. Rather than have the notebook designer decide ahead of time
which versions should be preserved, we embrace \emph{transparent} versioning.
This is similar to models like Jankun-Kelly et al.'s p-set calculus \cite{Jankun-Kelly:2007:MFV}
and VisTrails's version tree \cite{Callahan:2006:VVM}, where every change in the state of the system is tracked.

Implementing the entire version-control subsystem would be a large
undertaking. Fortunately, we can build RCloud on top of Github's
\emph{gists}~\cite{Github:2014:GG}. Github offers a HTTP interface
for creating simplified git repositories (the main limitation being the
use of text-only files in a single directory). The GitHub web-service
API provides most of the semantics we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.

Using GitHub was mainly a tactical decision to reduce our
time-to-prototype. At the same time, it had the beneficial effect of
forcing our backend to communicate with separate processes over HTTP
and JSON. This had the serendipitous advantage of quickly enabling
full search over RCloud notebooks (using Apache SOLR).
%
This perspective has implications for
systems-based research and prototyping in visual analytics.
%
In order for small teams to efficiently explore ideas in the present
software ecosystem, it pays off to use as many existing solutions as
possible.
%
Even though it involved learning a variety of standards that are not
directly related to visual analytics, our early decision to use the
GitHub gist API has been quite fruitful. We expect related projects
can gain similar benefits by adopting the same strategy.

>>>>

As a result, HTTP is the lingua franca of interprocess communication
(IPC). One of the design goals for RCloud was to provide a productive
environment for creators of data-analysis scripts, that also behaves
as a first-class citizen in the pre-existing ecosystem of computer
services and networks in an organization.

@@

One of the main differences between RCloud and other systems is the
notion that notebooks are ``always deployed''. This improves the
ability to share and modify experiments and compose the results
without binding to specific versions of a notebook.
On the other hand, there are a variety of situations where it might
be necessary to name a specific version. Rather than force notebook
designers to decide which versions need to be preserved,
we embrace \emph{transparent} versioning.
This is similar to models like Jankun-Kelly et al.'s p-set calculus \cite{Jankun-Kelly:2007:MFV}
and VisTrails's version tree \cite{Callahan:2006:VVM}, where every change in the state of the system is tracked.

Implementing an entire version-control subsystem from scratch would
be a large undertaking. Fortunately, we can build RCloud on top of Github's
\emph{gists}~\cite{Github:2014:GG}. Github offers a HTTP interface
for creating simplified git repositories (the main limitation being the
use of text-only files in a single directory). The GitHub web-service
API provides most of the semantics we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.

Using GitHub was mainly a tactical decision to reduce our
time-to-prototype. At the same time, it had the beneficial effect of
allowing our backend to communicate with separate processes over HTTP
and JSON. This had the serendipitous advantage of quickly enabling
full search over RCloud notebooks, using Apache SOLR.
%
This perspective has implications for
systems-based research and prototyping in visual analytics:
For small teams to efficiently explore novel ideas toward the
improvement of a large, complex software ecosystem, it pays off
to use existing solutions as far as possible.
%
Even though it involved learning standards that are not directly
related to visual analytics, our early decision to use the GitHub
gist API was succesful. We expect related projects would gain many
of the same benefits by adopting a similar strategy.
